Discrete Quantum Theories and Computing

### 2019-02-06 dissertation.tex
dissertation.lyx Fixing the space in tables...

\noindent \begin{centering}
\caption{\label{tab:committee}I want to thank the professors co-authored with

@@

\par\end{centering}
\noindent \centering{}\noindent \centering{}%
\begin{tabularx}{\linewidth}{M{MATH-M781}M{Co-author}M{(\textarabic{عمرو صبري})}M{Co-author}M{Thurston}M{Shouhong}M{Lawrence}Y}

@@

 & Co-author & Co-author & Co-author &  &  &  & Most of Chapters~\ref{chap:Conventional-Quantum-Theory}, \ref{chap:Unrestricted-Finite-Fields},
and~\ref{chap:DQT=000026DQC} is based on~\cite{geometry2013} and~\cite{DQT2014}; most of Chapter~\ref{chap:QIVPM} is
based on~\cite{THOS2018}.\tabularnewline\addlinespace

@@

Research committee & Co-chair & Co-chair & Member & Co-chair & Member &  & Some questions in the dissertation proposal are answered in Chapter~\ref{chap:QIVPM},
and some of the remaining ones are explained in Chapter~\ref{chap:Further-Discussion}.\tabularnewline\addlinespace

@@

PHYS-P700 & Instructor &  &  &  &  &  & The textbook~\cite{Mermin2007} is
cited in Chapters~\ref{chap:Introduction}, \ref{chap:Conventional-Quantum-Theory},

@@

MATH-M781 &  &  &  &  &  & Instructor & The presented paper~\cite{Abramsky2012} is cited in Sec.~\ref{subsec:Classical-and-Quantum}.\tabularnewline\addlinespace

@@

Tier 3 committee & Co-chair & Co-chair &  & Member &  & Co-chair & Two of
the assigned papers in the Tier 3 exam, \cite{Schumacher2012-SCHMQT}
and~\cite{doi:10.1142/S0217984913500644},

@@

@@

\end{tabularx}

@@

\begin{sidewaystable}
\noindent \begin{centering}
\caption{\label{tab:teacher}The knowledge I learned from teachers in other
courses help me consolidate this dissertation, where IUB stands for
Indiana University Bloomington and NTU stands for National Taiwan
University.}
\par\end{centering}
%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{ccM{MATH-M522}M{English~for~International~Graduate~Students:}Y}
\toprule 
\addlinespace
Instructor & School & Course & Title & Relation with the dissertation\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
Tom Lewis & IUB & SLST-T501 & English for International Graduate Students:\linebreak[3] Academic
Writing & Sec.~\ref{subsec:Discrete-Deutsch-algorithm} is based on the term
paper.\tabularnewline\addlinespace
\midrule 
\addlinespace
Michael A. Mandell & IUB & MATH-M522 & Topology II & The textbook~\cite{Hatcher2001} is cited in Chapters~\ref{chap:Introduction}
and~\ref{chap:Conventional-Quantum-Theory}.\tabularnewline\addlinespace
\midrule 
\addlinespace
Steven Myers & IUB & CSCI-B502 & Computational Complexity & The textbook~\cite{AroraBarak2009} is cited in Sec.~\ref{modalquantumcomputing}.\tabularnewline\addlinespace
\midrule 
\addlinespace
Valery Lunts & IUB & MATH-M501 & Survey of Algebra & The textbook~\cite{DummitFoote2004} is cited in Sec.~\ref{sec:background},
\ref{discretequantumtheoryI}, and~\ref{sec:Discrete-Quantum-Theory-(II)}.\tabularnewline\addlinespace
\midrule 
\addlinespace
Huah Chu (朱樺\textenglish[variant=american]{)} & NTU & 221~U3830\linebreak[3] 221~U3840 & Algebra(I)\linebreak[3] Algebra(II) & The textbook~\cite{Artin1991} is cited in Sec.~\ref{CQC1qubitBloch.sec},
\ref{sec:background}, \ref{discretequantumtheoryI}, and~\ref{sec:Discrete-Quantum-Theory-(II)}.\tabularnewline\addlinespace
\midrule
\addlinespace
Jin-Tzu Chen (陳金次) & NTU & 201~31300 & Functions of a Complex Variable & The textbook~\cite{GAMELIN2003} is cited in Sec.~\ref{subsec:Vector-Spaces}.\tabularnewline\addlinespace
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{sidewaystable}
To have the current the content of the dissertation, I would also
like to thank John Gardiner for some inspiring discussion about quantum
probability measures over finite fields~\cite{Gardiner2014}. His
results were improved and motivated the transition from DQT to QIVPM
in Sec.~\ref{sec:Toward-IVPM}. Besides of my committee members,
the knowledge I learned from teachers in other courses help me consolidate
this dissertation as listed in Table~\ref{tab:teacher}. Although
not directly relating to the dissertation, I have to thank for your
teaching to make me more mature and easier to communicate with others,
especially my English teachers, Traci Nagle, Elizabeth ``Betsy''
Merceron, and Kexin ``Casey'' Chen (谌可心).

@@

\begin{sidewaystable}
\caption{\label{tab:Comparison}Comparison between the classical quantities
and their mathematical models with the inspired quantum models.}

\centering{}\centering{}%
\begin{tabularx}{\linewidth}{YcYY}
\toprule 
\addlinespace
Physical quantity & Discrete & Discrete with extreme small units & Continuous, or no evidence to support they are discrete\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational device & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Mathematical representation & Integer & Real number & Real number\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational model  & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How does the theorical model predict the behavior of the physical device? & 
Reliably & 
Not reliably, but after we have better technology to manipulate the minimum units, we can more reliably predict the physical phenomenon and their computational power by a better discrete model. & 
Not reliably, because precision can never be high enough, and the difference between their computational power might not be able compensated by error analysis techniques. \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Inspired quantum model &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and two
possibilities for classical quantities inspires two types of quantum
models. Like its classical counterparts, quantum circuit model, the
most widely used model of quantum computing, manipulates quantities
which are not discrete apparently: the probability amplitudes of quantum
states, which are assumed in the field of complex numbers in the ``conventional
quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.}. If the probability amplitudes are assumed to be ultimately discrete,
we choose to replace the field of complex numbers by discrete finite
fields for two reasons. On one hand, CQT is built upon the operations
among the probability amplitudes, and the probability amplitudes have
the required operations since the complex numbers form a field. On
the other hand, based on finite fields, we will define the fraction-like
cardinal probability which has extremely small units when the size
of the field is extremely large. Alternatively, even if the probability
amplitudes might never be found quantized, considering the precision
of quantities could still result in a better model. However, comparing
to the precision of probability amplitudes, it is easier to consider
the precision of their inducing probabilities because the idea of
``imprecise probability'' has well-studied classically \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
\begin{comment}
\todo{Make sure whether realistic quantum computers could branch
their computation based on their probability amplitudes or measured
probabilities? If they can only branch on their measured probabilities,
it will provide another reason why we want to study the precision
of the inducing probabilities\ldots{} }
\end{comment}
This idea will be extended to quantum domain as quantum interval-valued
probability measure (QIVPM).

@@

\begin{tabularx}{\linewidth}{YYYYYYY}
\toprule 
\addlinespace
 & Conventional & Modal & Discrete (I) & Discrete (II) & QPMFF & QIVPM\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 & Chapter~\ref{chap:Conventional-Quantum-Theory} & Sec.~\ref{modalquantum} & Sec.~\ref{discretequantumtheoryI} to~\ref{sec:Geometry-of-Entangled-States} & Sec.~\ref{sec:Discrete-Quantum-Theory-(II)} & Sec.~\ref{sec:Toward-IVPM} & Chapter~\ref{chap:QIVPM}\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
States space & $\mathbb{C}^{D}$ & $\Fq^{D}$ & $\Fpp^{D}$ & Local region in $\Fpp^{D}$ & $\Fpp^{D}$ & $\mathbb{C}^{D}$\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Likelihood of events is predicted by & Real-valued probability & Possible or impossible & Possible or impossible & Cardinal probability & Real-valued probability, but no sensible Born rule & Interval-valued probability\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Expectation value & Defined & Undefined & Formally defined & Undefined &  & Defined\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 &  & Sec.~\ref{modalquantumcomputing} & Sec.~\ref{discretequantumcomputingI} & Sec.~\ref{discretequantumcomputingII} &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Deutsch's algorithm & Yes & Maybe no & Yes & Yes &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Efficiently solve $\usat$ & Unlikely & Yes & Partially & Unlikely &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Grover search algorithm & Yes &  &  & Yes &  & \tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The organization of this thesis is summarized in Table~\ref{tab:organization},
and finally, Chapter~\ref{chap:Further-Discussion} contains further
extended questions, especially, briefly explaining why it might be
hard to recover Gleason's theorem for general QIVPMs.

@@

As an example, consider a three-dimensional Hilbert space with orthonormal
basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $ and the observable
$\mathbf{O}_{0}$ defined in Eq.~(\ref{eq:observable}). Two fragments
of valid probability measures~$\mu_{1}$ and $\mu_{2}$ that can
be associated with this space are defined in Table~\ref{tab:quantum-probability-measure}.
By the Born rule, the first probability measure corresponds to the
quantum system being in the pure state $\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}$
and the second corresponds to the quantum system being in the state
$\frac{\proj{0}+\proj{2}}{2}$. The expectation values of the observable
$\mathbf{O}$, $\expval{\mathbf{O}}_{\mu_{1,2}}$, are $1.5$ in the
first case and $2$ in the second. The quantum expectation value can
also be used to decide whether a state is entangled or not for multipartite
systems as we describe in the following section.
\begin{table}
\noindent \begin{centering}
\caption{\label{tab:quantum-probability-measure}Two fragments of valid probability
measures~$\mu_{1}$ and $\mu_{2}$.}
\par\end{centering}
\noindent \centering{}%
\begin{tabular}{ccccccccccc}
\toprule 
\addlinespace
$\ket{\Psi}$ & $\ket{0}$ & $\ket{1}$ & $\ket{2}$ & $\frac{\ket{0}+\ket{1}}{\sqrt{2}}$ & $\frac{\ket{0}+\rmi\ket{1}}{\sqrt{2}}$ & $\frac{\ket{0}+\ket{2}}{\sqrt{2}}$ & $\frac{\ket{0}+\rmi\ket{2}}{\sqrt{2}}$ & $\frac{\ket{1}+\ket{2}}{\sqrt{2}}$ & $\frac{\ket{1}+\rmi\ket{2}}{\sqrt{2}}$ & \ldots\tabularnewline\addlinespace
\midrule
\addlinespace
\addlinespace
$\mu_{1}\left(\proj{\Psi}\right)$ & $\frac{1}{2}$ & $\frac{1}{2}$ & $0$ & $1$  & $\frac{1}{2}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & \ldots\tabularnewline\addlinespace
\addlinespace
\addlinespace
$\mu_{2}\left(\proj{\Psi}\right)$ & $\frac{1}{2}$ & $0$ & $\frac{1}{2}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & $\frac{1}{2}$ & $\frac{1}{2}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & \ldots\tabularnewline\addlinespace
\bottomrule
\addlinespace
\end{tabular}
\end{table}

@@

\centering{}%
\begin{tabular}{ccccccccc}
\toprule 
\addlinespace
$j$ & $0$ & $1$ & $2$ & $3$ & $4$ & $5$ & $6$ & $7$\tabularnewline\addlinespace
\midrule
\addlinespace
\addlinespace
$\left(1+\rmi\right)^{j}$ & $1$ & $1+\rmi$ & $-\rmi$ & $1-\rmi$ & $-1$ & $-1-\rmi$ & $\rmi$ & $-1+\rmi$\tabularnewline\addlinespace
\addlinespace
\addlinespace
$\left(1-\rmi\right)^{j}$ & $1$ & $1-\rmi$ & $\rmi$ & $1+\rmi$ & $-1$ & $-1+\rmi$ & $-\rmi$ & $-1-\rmi$\tabularnewline\addlinespace
\addlinespace
\addlinespace
$\left(-1+\rmi\right)^{j}$ & $1$ & $-1+\rmi$ & $\rmi$ & $-1-\rmi$ & $-1$ & $1-\rmi$ & $-\rmi$ & $1+\rmi$\tabularnewline\addlinespace
\addlinespace
\addlinespace
$\left(-1-\rmi\right)^{j}$ & $1$ & $-1-\rmi$ & $-\rmi$ & $-1+\rmi$ & $-1$ & $1+\rmi$ & $\rmi$ & $1-\rmi$\tabularnewline\addlinespace
\bottomrule
\addlinespace
\end{tabular}
\end{table}

@@

\centering{}%
\begin{tabular}{cccc}
\toprule 
\addlinespace
$c=g^{\left(p+1\right)k}$ & $\left(p+1\right)k$ & $k$ & $\fnorm*{g^{\left(p+1\right)k}}=g^{k}$\tabularnewline\addlinespace
\midrule
\addlinespace
\addlinespace
$-1$ & $4$ & $1$ & $1-\rmi$\tabularnewline\addlinespace
\addlinespace
\addlinespace
$1$ & $0$ & $0$ & $1$\tabularnewline\addlinespace
\bottomrule
\addlinespace
\end{tabular}
\end{table}

@@

For a single qubit Deutsch problem, the probability amplitudes are
between $-2$ and $2$, and the dimension $D=2^{1+1}=4$, so we want
to have 
\begin{equation}
k\geq2^{3\cdot1+2}+1=2^{5}+1=33\,.
\end{equation}
The least prime satisfying the above condition is $k=37$, and thus
$\pi\left(37\right)=12$ and $p=422231$, as shown in the extended
elements Table~\ref{table_long}. 
\begin{table}
\noindent \centering{}\caption{\label{table_long}Extension of transitively ordered elements.}
\begin{tabular}{cccccccccc}
\toprule 
\addlinespace
$p$ & \ldots{} & $422231$ & \ldots{} & $196265095009$ & \ldots{} &  & \ldots{} &  & \ldots\tabularnewline\addlinespace
\addlinespace
\midrule 
\addlinespace
$k$ & \ldots{} & ${\bf 37}$ & \ldots{} & ${\bf 131}$ & \ldots{} & ${\bf 257}$ & \ldots{} & ${\bf 32771}$ & \ldots\tabularnewline\addlinespace
\midrule
\addlinespace
\addlinespace
$\pi\left(k\right)$ & \ldots{} & $12$ & \ldots{} & $32$ & \ldots{} & $55$ & \ldots{} & $3513$ & \ldots\tabularnewline\addlinespace
\bottomrule
\addlinespace
\end{tabular}
\end{table}

>>>>

\noindent \centering{}\caption{\label{tab:committee}I want to thank the professors co-authored with

@@

%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{M{MATH-M781}M{Co-author}M{(\textarabic{عمرو~صبري})}M{Co-author}M{Thurston}M{Shouhong}M{Lawrence}Y}

@@

 & Co-author & Co-author & Co-author &  &  &  & Most of Chapters~\ref{chap:Conventional-Quantum-Theory}, \ref{chap:Unrestricted-Finite-Fields},and~\ref{chap:DQT=000026DQC}
is based on~\cite{geometry2013} and~\cite{DQT2014}; most of Chapter~\ref{chap:QIVPM}
is based on~\cite{THOS2018}.\tabularnewline\addlinespace

@@

Research committee & Co-chair & Co-chair & Member & Co-chair & Member &  & Some questions in the dissertation proposal are answered in Chapter~\ref{chap:QIVPM},and
some of the remaining ones are explained in Chapter~\ref{chap:Further-Discussion}.\tabularnewline\addlinespace

@@

PHYS-P700 & Instructor &  &  &  &  &  & The textbook~\cite{Mermin2007} of \emph{Quantum Computation and Information}
is cited in Chapters~\ref{chap:Introduction}, \ref{chap:Conventional-Quantum-Theory},

@@

CSCI-B629 &  & Instructor &  &  &  &  & The textbook~\cite{hottbook2013} of \emph{Homotopy Type Theory}
is cited in Sec.~\ref{discretequantumtheoryIIb}.\tabularnewline\addlinespace

@@

Tier 3 committee & Co-chair & Co-chair &  & Member &  & Co-chair & Two of the assigned papers in the Tier 3 exam \cite{Schumacher2012-SCHMQT,doi:10.1142/S0217984913500644}

@@

MATH-M781 &  &  &  &  &  & Instructor & The presented paper~\cite{Abramsky2012} of \emph{Coalgebra} is
cited in Sec.~\ref{subsec:Classical-and-Quantum}.\tabularnewline\addlinespace

@@

\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 

@@

To have the current the content of the dissertation, I would also
like to thank John Gardiner for some inspiring discussion about quantum
probability measures over finite fields~\cite{Gardiner2014}. His
results were improved and motivated the transition from DQT to QIVPM
in Sec.~\ref{sec:Toward-IVPM}. Besides of my committee members,
the knowledge I learned from teachers in other courses help me consolidate
this dissertation as listed in Table~\ref{tab:teacher}. Although
not directly relating to the dissertation, I have to thank for your
teaching to make me more mature and easier to communicate with others,
especially my English teachers, Traci Nagle, Elizabeth ``Betsy''
Merceron, and Kexin ``Casey'' Chen (谌可心).
\begin{table}[t]
\noindent \begin{centering}
\caption{\label{tab:teacher}The knowledge I learned from teachers in other
courses help me consolidate this dissertation, where IUB stands for
Indiana University Bloomington and NTU stands for National Taiwan
University.}
\par\end{centering}
%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{M{Jin-Tzu~Chen}cM{MATH-M522}Y}
\toprule 
\addlinespace
Instructor & School & Course & Relation with the dissertation\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
Tom Lewis & IUB & SLST-T501 & Sec.~\ref{subsec:Discrete-Deutsch-algorithm} is based on the term
paper of \emph{Academic Writing}.\tabularnewline\addlinespace
\midrule 
\addlinespace
Michael\nolinebreak[3] A. Mandell & IUB & MATH-M522 & The textbook~\cite{Hatcher2001} of \emph{Topology II} is cited
in Chapters~\ref{chap:Introduction} and~\ref{chap:Conventional-Quantum-Theory}.\tabularnewline\addlinespace
\midrule 
\addlinespace
Steven Myers & IUB & CSCI-B502 & The textbook~\cite{AroraBarak2009} of \emph{Computational Complexity}
is cited in Sec.~\ref{modalquantumcomputing}.\tabularnewline\addlinespace
\midrule 
\addlinespace
Valery Lunts & IUB & MATH-M501 & The textbook~\cite{DummitFoote2004} of \emph{Survey of Algebra}
is cited in Sec.~\ref{sec:background}, \ref{discretequantumtheoryI},
and~\ref{sec:Discrete-Quantum-Theory-(II)}.\tabularnewline\addlinespace
\midrule 
\addlinespace
Huah Chu\linebreak[3] (朱樺\textenglish[variant=american]{)} & NTU & 221~U3830\linebreak[3] 221~U3840 & The textbook~\cite{Artin1991} of \emph{Algebra (I)} and \emph{(II)}
is cited in Sec.~\ref{CQC1qubitBloch.sec}, \ref{sec:background},
\ref{discretequantumtheoryI}, and~\ref{sec:Discrete-Quantum-Theory-(II)}.\tabularnewline\addlinespace
\midrule
\addlinespace
Jin-Tzu Chen\linebreak[3] (陳金次) & NTU & 201~31300 & The textbook~\cite{GAMELIN2003} of \emph{Functions of a Complex Variable}
is cited in Sec.~\ref{subsec:Vector-Spaces}.\tabularnewline\addlinespace
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{table}

@@

\begin{sidewaystable}
\caption{\label{tab:Comparison}Comparison between the classical quantities
and their mathematical models with the inspired quantum models.}
%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{YcYY}
\toprule 
\addlinespace
Physical quantity & Discrete & Discrete with extreme small units & Continuous, or no evidence to support they are discrete\tabularnewline\addlinespace
\midrule 
\addlinespace
Computational device & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
Mathematical representation & Integer & Real number & Real number\tabularnewline\addlinespace
\midrule 
\addlinespace
Computational model & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
How does the theoretical model predict the behavior of the physical
device? & Reliably & Not reliably, but after we have better technology to manipulate the
minimum units, we can more reliably predict the physical phenomenon
and their computational power by a better discrete model. & Not reliably, because precision can never be high enough, and the
difference between their computational power might not be able compensated
by error analysis techniques.\tabularnewline\addlinespace
\midrule 
\addlinespace
Inspired quantum model &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline\addlinespace
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and two
possibilities for classical quantities inspires two types of quantum
models. Like its classical counterparts, quantum circuit model, the
most widely used model of quantum computing, manipulates quantities
which are not discrete apparently: the probability amplitudes of quantum
states, which are assumed in the field of complex numbers in the ``conventional
quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.}. If the probability amplitudes are assumed to be ultimately discrete,
we choose to replace the field of complex numbers by discrete finite
fields for two reasons. On one hand, CQT is built upon the operations
among the probability amplitudes, and the probability amplitudes have
the required operations since the complex numbers form a field. On
the other hand, based on finite fields, we will define the fraction-like
cardinal probability which has extremely small units when the size
of the field is extremely large. Alternatively, even if the probability
amplitudes might never be found quantized, considering the precision
of quantities could still result in a better model. However, comparing
to the precision of probability amplitudes, it is easier to consider
the precision of their inducing probabilities because the idea of
``imprecise probability'' has well-studied classically \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
\begin{comment}
\todo{Make sure whether realistic quantum computers could branch
their computation based on their probability amplitudes or measured
probabilities? If they can only branch on their measured probabilities,
it will provide another reason why we want to study the precision
of the inducing probabilities\ldots{} }
\end{comment}
This idea will be extended to quantum domain as quantum interval-valued
probability measure (QIVPM).

@@

%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{YYYYYYY}
\toprule 
\addlinespace
 & Conventional & Modal & Discrete (I) & Discrete (II) & QPMFF & QIVPM\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
 & Chapter~\ref{chap:Conventional-Quantum-Theory} & Sec.~\ref{modalquantum} & Sec.~\ref{discretequantumtheoryI} to \ref{sec:Geometry-of-Entangled-States} & Sec.~\ref{sec:Discrete-Quantum-Theory-(II)} & Sec.~\ref{sec:Toward-IVPM} & Chapter~\ref{chap:QIVPM}\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
States space & $\mathbb{C}^{D}$ & $\Fq^{D}$ & $\Fpp^{D}$ & Local region in $\Fpp^{D}$ & $\Fpp^{D}$ & $\mathbb{C}^{D}$\tabularnewline\addlinespace
\midrule 
\addlinespace
Likelihood of events is predicted by & Real-valued probability & Possible or impossible & Possible or impossible & Cardinal probability & Real-valued probability, but no sensible Born rule & Interval-valued probability\tabularnewline\addlinespace
\midrule 
\addlinespace
Expectation value & Defined & Undefined & Formally defined & Undefined &  & Defined\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
 &  & Sec.~\ref{modalquantumcomputing} & Sec.~\ref{discretequantumcomputingI} & Sec.~\ref{discretequantumcomputingII} &  & \tabularnewline\addlinespace
\midrule 
\addlinespace
Deutsch's algorithm & Yes & Maybe no & Yes & Yes &  & \tabularnewline\addlinespace
\midrule 
\addlinespace
Efficiently solve $\usat$ & Unlikely & Yes & Partially & Unlikely &  & \tabularnewline\addlinespace
\midrule 
\addlinespace
Grover search algorithm & Yes &  &  & Yes &  & \tabularnewline\addlinespace
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{sidewaystable}
The organization of this thesis is summarized in Table~\ref{tab:organization},
and finally, Chapter~\ref{chap:Further-Discussion} contains further
extended questions, especially, briefly explaining why it might be
hard to recover Gleason's theorem for general QIVPMs.

@@

As an example, consider a three-dimensional Hilbert space with orthonormal
basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $ and the observable
$\mathbf{O}_{0}$ defined in Eq.~(\ref{eq:observable}). Two fragments
of valid probability measures~$\mu_{1}$ and $\mu_{2}$ that can
be associated with this space are defined in Table~\ref{tab:quantum-probability-measure}.
By the Born rule, the first probability measure corresponds to the
quantum system being in the pure state $\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}$
and the second corresponds to the quantum system being in the state
$\frac{\proj{0}+\proj{2}}{2}$. The expectation values of the observable
$\mathbf{O}$, $\expval{\mathbf{O}}_{\mu_{1,2}}$, are $1.5$ in the
first case and $2$ in the second. The quantum expectation value can
also be used to decide whether a state is entangled or not for multipartite
systems as we describe in the following section.
\begin{table}
\noindent \begin{centering}
\caption{\label{tab:quantum-probability-measure}Two fragments of valid probability
measures~$\mu_{1}$ and $\mu_{2}$.}
\par\end{centering}
\noindent \centering{}%
\begin{tabular}{ccccccccccc}
\toprule 
\addlinespace
$\ket{\Psi}$ & $\ket{0}$ & $\ket{1}$ & $\ket{2}$ & $\frac{\ket{0}+\ket{1}}{\sqrt{2}}$ & $\frac{\ket{0}+\rmi\ket{1}}{\sqrt{2}}$ & $\frac{\ket{0}+\ket{2}}{\sqrt{2}}$ & $\frac{\ket{0}+\rmi\ket{2}}{\sqrt{2}}$ & $\frac{\ket{1}+\ket{2}}{\sqrt{2}}$ & $\frac{\ket{1}+\rmi\ket{2}}{\sqrt{2}}$ & \ldots\tabularnewline\addlinespace
\midrule
\addlinespace
$\mu_{1}\left(\proj{\Psi}\right)$ & $\frac{1}{2}$ & $\frac{1}{2}$ & $0$ & $1$  & $\frac{1}{2}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & \ldots\tabularnewline\addlinespace
\addlinespace
$\mu_{2}\left(\proj{\Psi}\right)$ & $\frac{1}{2}$ & $0$ & $\frac{1}{2}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & $\frac{1}{2}$ & $\frac{1}{2}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & \ldots\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}

@@

\centering{}%
\begin{tabular}{ccccccccc}
\toprule 
\addlinespace
$j$ & $0$ & $1$ & $2$ & $3$ & $4$ & $5$ & $6$ & $7$\tabularnewline\addlinespace
\midrule
\addlinespace
$\left(1+\rmi\right)^{j}$ & $1$ & $1+\rmi$ & $-\rmi$ & $1-\rmi$ & $-1$ & $-1-\rmi$ & $\rmi$ & $-1+\rmi$\tabularnewline\addlinespace
\addlinespace
$\left(1-\rmi\right)^{j}$ & $1$ & $1-\rmi$ & $\rmi$ & $1+\rmi$ & $-1$ & $-1+\rmi$ & $-\rmi$ & $-1-\rmi$\tabularnewline\addlinespace
\addlinespace
$\left(-1+\rmi\right)^{j}$ & $1$ & $-1+\rmi$ & $\rmi$ & $-1-\rmi$ & $-1$ & $1-\rmi$ & $-\rmi$ & $1+\rmi$\tabularnewline\addlinespace
\addlinespace
$\left(-1-\rmi\right)^{j}$ & $1$ & $-1-\rmi$ & $-\rmi$ & $-1+\rmi$ & $-1$ & $1+\rmi$ & $\rmi$ & $1-\rmi$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}

@@

\centering{}%
\begin{tabular}{cccc}
\toprule 
\addlinespace
$c=g^{\left(p+1\right)k}$ & $\left(p+1\right)k$ & $k$ & $\fnorm*{g^{\left(p+1\right)k}}=g^{k}$\tabularnewline\addlinespace
\midrule
\addlinespace
$-1$ & $4$ & $1$ & $1-\rmi$\tabularnewline\addlinespace
\addlinespace
$1$ & $0$ & $0$ & $1$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}

@@

For a single qubit Deutsch problem, the probability amplitudes are
between $-2$ and $2$, and the dimension $D=2^{1+1}=4$, so we want
to have 
\begin{equation}
k\geq2^{3\cdot1+2}+1=2^{5}+1=33\,.
\end{equation}
The least prime satisfying the above condition is $k=37$, and thus
$\pi\left(37\right)=12$ and $p=422231$, as shown in the extended
elements Table~\ref{table_long}. 
\begin{table}
\noindent \centering{}\caption{\label{table_long}Extension of transitively ordered elements.}
\begin{tabular}{cccccccccc}
\toprule 
\addlinespace
$p$ & \ldots{} & $422231$ & \ldots{} & $196265095009$ & \ldots{} &  & \ldots{} &  & \ldots\tabularnewline\addlinespace
\midrule 
\addlinespace
$k$ & \ldots{} & ${\bf 37}$ & \ldots{} & ${\bf 131}$ & \ldots{} & ${\bf 257}$ & \ldots{} & ${\bf 32771}$ & \ldots\tabularnewline\addlinespace
\midrule
\addlinespace
$\pi\left(k\right)$ & \ldots{} & $12$ & \ldots{} & $32$ & \ldots{} & $55$ & \ldots{} & $3513$ & \ldots\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
